---
title: "GO enrichment analysis and simulation studies"
author: "Greg Finak and Masanao Yajima"
date: "September 3, 2015"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Additional Analyses}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

We look at GO enrichment of DE genes using MAST, with and without CDR and compare it to GO enrichment of DE genes using other RNASeq tools. 

We compare the ratio of immune specific to non-specific enriched modules.

We'll use GO.db and org.Hs.eg.db for the MAIT cells.

We'll use the the hypergeometric for testing enrichment. 

We load the results and databases.

```{r}
library(data.table)
library(plyr)
library(GO.db)
library(stringr)
library(grid)
library(gridExtra)
library(ggplot2)
library(mHG)
library(org.Hs.eg.db)
library(org.Mm.eg.db)
#diffex = readRDS("../inst/extdata/method_comp_logfc_qvalue.rds")
diffex = readRDS("../inst/extdata/comb_fdr_logfc.rds")

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}
```


```{r}
goterms = select(org.Hs.eg.db,diffex[,primerid],"GO","ENTREZID")
setDT(goterms)
goterms = goterms[ONTOLOGY%in%"BP"]
#only experimental evidence
goterms = goterms[EVIDENCE%in%c("EXP","IDA","IPI","IMP","IGI","IEP")]
unique.go = goterms[,unique(ENTREZID),.(GO)]
terms = select(GO.db,columns = c("GOID","TERM"),unique(goterms$GO))
setnames(unique.go,"GO","GOID")
setnames(unique.go,"V1","ENTREZID")
setDT(terms)
terms = unique(terms)
unique.go = unique(unique.go)
unique.go = merge(unique.go,terms,by="GOID")
unique.go[,N:=length(ENTREZID),.(GOID)]
offspring = GOBPOFFSPRING[["GO:0002376"]]
```

```{r}
setnames(diffex,"primerid","ENTREZID")
thresholds = c(0.01,0.05,0.1)
prob_thresh=0.01
NN = length(unique(unique.go$ENTREZID))
methods = c("zlm_fdr","zlm_ng_fdr","zlm_fix_fdr","zlm_fix_ng_fdr","limma_fdr","limma_fdr","limma_ng_fdr","deseq_fdr","deseq_ng_fdr","edger_fdr","edger_ng_fdr","scde_fdr")
loops = expand.grid(thresholds,methods)
thresh=0.05

meth="zlm_ng_fdr"
results = apply(loops,1,function(looprow){
  thresh = looprow[1]
  meth = as.character(looprow[2])
  fc = gsub("fdr","log2fc",meth)
  res = unique.go[ENTREZID%in%diffex[get(meth)<thresh&abs(get(fc))>log2(1.25),ENTREZID]][,.(prob = 1-phyper(q=length(ENTREZID),m = N, n = NN-N,k = length(diffex[get(meth)<thresh,ENTREZID])),TERM,N,hit = length(ENTREZID)),.(GOID)]
 nr.imm = length(unique(res)[GOID%in%offspring&p.adjust(prob,"fdr")<prob_thresh,GOID]) 
nr.tot = length(unique(res)[p.adjust(prob,"fdr")<prob_thresh,GOID])
data.frame(thresh,meth,nr.imm,nr.tot,nr.imm/nr.tot)
})
```



## Plot of the rate of detection of enriched immune GO terms vs FDR threshold for DE genes across different methods.

```{r}
results_r = ldply(results)
setDT(results_r)
results_r[,meth:=gsub("zlm_fix","zlmfix",meth)]
results_r[,rate := nr.imm/nr.tot]
results_r[,thresh:=as.numeric(as.character(thresh))]
results_r[,method_global := str_split_fixed(meth,"_",3)[,1]]
results_r[,withng := factor(str_split_fixed(meth,"_",3)[,2],labels=c("No CDR control","CDR control"))]
p1 = ggplot(results_r[!(method_global%in%"zlmfix")])+aes(x=thresh,y=rate,color=method_global)+geom_point()+geom_line()+scale_x_continuous("FDR Threshold")+scale_y_continuous("Proportion of Immune\nSpecific GO Modules")+facet_wrap(~withng)+scale_color_brewer("Method",palette = 2,type="qual",limits=c("limma","zlm","deseq","edger","scde"),labels=c("Limma","MAST","DESeq","EdgeR","SCDE"))+theme_gray()
leg = g_legend(p1)
p1
```


## mDC
```{r,eval=FALSE}
diffex.mdc = readRDS("../inst/extdata/method_comp_logfc_qvalue_mDC.rds")
setnames(diffex.mdc,"primerid","SYMBOL")
diffex.mdc$SYMBOL=gsub("^(.)","\\U\\1",tolower(diffex.mdc$SYMBOL),perl=TRUE)
goterms = select(org.Mm.eg.db,diffex.mdc[,SYMBOL],"GO","SYMBOL")
setDT(goterms)
goterms = goterms[ONTOLOGY%in%"BP"]
#only experimental evidence
goterms = goterms[EVIDENCE%in%c("EXP","IDA","IPI","IMP","IGI","IEP")]
unique.go = goterms[,unique(SYMBOL),.(GO)]
terms = select(GO.db,columns = c("GOID","TERM"),goterms$GO)
setnames(unique.go,"GO","GOID")
setnames(unique.go,"V1","SYMBOL")
setDT(terms)
terms = unique(terms)
unique.go = unique(unique.go)
unique.go = merge(unique.go,terms,by="GOID")
unique.go[,N:=length(SYMBOL),.(GOID)]
NN=length(unique(unique.go[,SYMBOL]))
```


## GO enrichment using hypergeometric for mDC.
```{r,eval=FALSE}
thresholds = c(0.01,0.05,0.1)
prob_thresh=0.01
methods = c("fdr_lma_2_wo","fdr_lma_2_ng","fdr_zlm_2_wo","fdr_zlm_2_ng","fdr_lma_4_wo","fdr_lma_4_ng","fdr_zlm_4_wo","fdr_zlm_4_ng","fdr_lma_6_wo","fdr_lma_6_ng","fdr_zlm_6_wo","fdr_zlm_6_ng")
loops = expand.grid(thresholds,methods)
results.mdc = apply(loops,1,function(looprow){
  thresh = looprow[1]
  meth = as.character(looprow[2])
  fc = gsub("fdr","logFC",meth)
    res = unique.go[SYMBOL%in%diffex.mdc[get(meth)<thresh&abs(get(fc))>log2(1.25),SYMBOL]][,.(prob = 1-phyper(q=length(SYMBOL),m = N, n = NN-N,k = length(diffex.mdc[get(meth)<thresh,SYMBOL])),TERM,N,hit = length(SYMBOL)),.(GOID)]

nr.imm = length(unique(res)[GOID%in%offspring&p.adjust(prob,"fdr")<prob_thresh,GOID]) 

nr.tot = length(unique(res)[p.adjust(prob,"fdr")<prob_thresh,GOID])
data.frame(thresh,meth,nr.imm,nr.tot,nr.imm/nr.tot)
})
```

```{r,eval=FALSE}
results.mdc_r = ldply(results.mdc)
setDT(results.mdc_r)
results.mdc_r[,thresh:=as.numeric(as.character(thresh))]
results.mdc_r[,method_global := str_split_fixed(meth,"_",4)[,2]]
results.mdc_r[,time := str_split_fixed(meth,"_",4)[,3]]
results.mdc_r[,withng := factor(str_split_fixed(meth,"_",4)[,4],labels=c("No CDR control","CDR control"))]
results.mdc_r = results.mdc_r[,rate := mean(nr.imm)/(nr.tot),]
results.mdc_r=results.mdc_r[,.(rate = mean(rate)),.(withng,method_global,thresh)]
p2 = ggplot(results.mdc_r)+aes(x=thresh,y=rate,color=method_global)+geom_point()+geom_line()+scale_x_continuous("FDR Threshold")+scale_y_continuous("Proportion of Immune\nSpecific GO Modules")+facet_grid(~withng)+scale_color_brewer("Method",palette = 2,type="qual",limits=c("lma","zlm"),labels=c("Limma","MAST"))+theme_gray()
```

```{r,eval=FALSE,fig.width=8,fig.height=6}
library(grid)
p3 = grid.arrange(arrangeGrob(p1+theme(legend.position="none"),p2+theme(legend.position="none"),nrow=2),leg,nrow=1,widths=c(10,2))
pdf(file="../inst/extdata/output/SuppFig_10.pdf",width=6,height=8)
p3
grid.draw(p3)
grid.text(c("A","B"),x=c(0.01,0.01),y=c(0.95,0.55))
dev.off()
```

```{r sfig10a}
pdf(file="../inst/extdata/output/SuppFig_10_a.pdf",width=6,height=4)
p1
dev.off()
```

### GSEA of GO BP Modules in mDC

```{r,eval=FALSE}
library(RamiGO)
data("c5.go.mapping")
gsea_go = readRDS("../inst/extdata/gsea_result_mDC.rds")
gomap = melt(sapply(gsea_go$module,function(x)c5.go.mapping[match(x,c5.go.mapping[,1]),2]))
gomap = cbind(rownames(gomap),gomap)
colnames(gomap) =c("TERM","GOID")
setDT(gomap)
setnames(gsea_go,"module","TERM")
gsea_go = merge(gsea_go_ids,gsea_go,by="TERM")
setnames(gsea_go,gsub("zlm_fix","zlmfix",colnames(gsea_go)))
gsea_go = melt(gsea_go,id=c("TERM","GOID"))
gsea_go = cbind(gsea_go,data.table((gsea_go[,str_split_fixed(variable,"_",5)])))
setnames(gsea_go,c("V1","V2","V3","V4","V5"),c("V1","Time","signif","Method","ng"))
gsea_go[,V1:=NULL]
ggplot(ldply(lapply(c(0.01,0.05,0.1),function(x){
cbind(gsea_go[signif%in%"FDR",.(sum(GOID%in%offspring&value<substitute(eval(x)))/sum(value<substitute(eval(x) ))),.(Method,ng)],thresh=x)[,ng:=factor(ng,levels=c("","ng"),labels=c("No CDR","With CDR"))]
})))+aes(x=thresh,y=V1,color=Method)+facet_wrap(~ng)+geom_line()+geom_point()+scale_x_continuous("FDR Threshold")+scale_color_brewer("Method",palette = 2,type="qual",limits=c("limma","zlm","zlmfix"),labels=c("Limma","MAST","MAST (fixed threshold)"))+scale_y_continuous("Proportion of Enriched\nImmune Specific GO BP Modules")
```

### GSEA of GO BP Modules Enriched in MAIT

```{r,eval=FALSE}
gsea_go = readRDS("../inst/extdata/gsea_result.rds")
gomap = melt(sapply(gsea_go$module,function(x)c5.go.mapping[match(x,c5.go.mapping[,1]),2]))
gomap = cbind(rownames(gomap),gomap)
colnames(gomap) =c("TERM","GOID")
setDT(gomap)
setnames(gsea_go,"module","TERM")
gsea_go = merge(gsea_go_ids,gsea_go,by="TERM")
setnames(gsea_go,gsub("zlm_fix","zlmfix",colnames(gsea_go)))
gsea_go = melt(gsea_go,id=c("TERM","GOID"))
gsea_go = cbind(gsea_go,data.table((gsea_go[,str_split_fixed(variable,"_",3)])))
setnames(gsea_go,c("V1","V2","V3"),c("var","Method","ng"))
ggplot(ldply(lapply(c(0.01,0.05,0.1),function(x){
cbind(gsea_go[var%in%"FDR",.({foo=sum(GOID%in%offspring&value<substitute(eval(x)))/sum(value<substitute(eval(x) ));ifelse(is.nan(foo),0,foo)}),.(Method,ng)],thresh=x)[,ng:=factor(ng,levels=c("","ng"),labels=c("No CDR","With CDR"))]
})))+aes(x=thresh,y=V1,color=Method)+facet_wrap(~ng)+geom_line(alpha=0.7)+geom_point(alpha=0.7)+scale_x_continuous("FDR Threshold")+scale_color_brewer("Method",palette = 2,type="qual")+scale_y_continuous("Proportion of Enriched Immune\nSpecific GO BP Modules")
```

# Simulation study

```{r load_library}
library(abind)
library(arm)
library(data.table)
library(ggplot2)
library(grid)
library(devtools)
#install_github("RGlab/MAST")
library(MAST)
library(mvtnorm)
library(GSEABase)
library(limma)
library(MCMCpack)
library(pheatmap)
library(GGally)
library(RColorBrewer)
library(reshape2)
library(AnnotationDbi)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
```
```{r global_options, include=FALSE}
knitr::opts_chunk$set( fig.width = 14, 
                       echo      = FALSE, 
                       warning   = FALSE, 
                       message   = FALSE )
```

```{r define_constants,echo=FALSE, message=FALSE}
min_gene_in_module <- 5
FCTHRESHOLD        <- log2(1.5)
nrep           <- 10
options("mc.cores"=detectCores())
```
## For portability, all the time consuming steps are not evaluated.
## Please set ALL the eval flag in the knitr chunks to TRUE before running the code.

```{r data_loading, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE }
set.seed(123123)

#install.packages("MASTDataPackage",repos="http://gull.fhcrc.org")
library(MASTDataPackage)
data(MASTDataPackage)
sca_mait_original <- sca_mait
sca_mait          <- subset(sca_mait,nGeneOn>4000)
eid               <- select(TxDb.Hsapiens.UCSC.hg19.knownGene,
                           keys=fData(sca_mait)$entrez,
                           keytype="GENEID",
                           columns=c("GENEID","TXNAME"))
ueid              <- unique(na.omit(eid)$GENEID)
sca_mait          <- sca_mait[,fData(sca_mait)$entrez%in%ueid]

MAST:::layer(sca_mait)   <- "tpm"
sca_mait_filtered <- sca_mait
unlogged          <- 2^exprs(sca_mait)-1
tt <- thresholdSCRNACountMatrix( unlogged              ,
                                 nbins          = 20   , 
                                 min_per_bin    = 30   ,
                                 return_log     = TRUE
                               )
#par(mfrow=c(6,5))
#plot(tt)

fd_filt                   <- fData(sca_mait)
fd_filt$cutpoint          <- tt$cutpoint[as.character(tt$bin)]

esetAll<-abind(sca_mait@.Data, atpm=tt$counts_threshold, along=3)

sca_mait_filtered<-FromMatrix('SingleCellAssay',esetAll,
                                cData(sca_mait_filtered),fd_filt)
dimnames(sca_mait_filtered)[[3]]<-dimnames(esetAll)[[3]]
MAST:::layer(sca_mait_filtered)<-"atpm"
sca_mait_filtered         <- sca_mait_filtered[,freq(sca_mait_filtered)>0.2]
sca                       <- sca_mait_filtered
MAST:::layer(sca)<-"atpm"
```


### fit ZLM

```{r fitzlm, eval=FALSE, echo=FALSE}
MAST:::layer(sca)  <- 'atpm'
# without ngeneson
cData(sca)$condition<-factor(cData(sca)$condition,levels=c("Unstim","Stim"))
zlm_cond <- zlm.SingleCellAssay( ~ condition , sca,  
                                 method='bayesglm', ebayes = TRUE, 
                                 ebayesControl =list(method = "MLE", model = "H1") ,  
                                 hook=deviance_residuals_hook )
lrt_cond  <- lrTest( zlm_cond,CoefficientHypothesis("conditionStim"))
wald_cond <- waldTest(zlm_cond,CoefficientHypothesis("conditionStim"))
logFC_no  <- getLogFC(zlm_cond,c(1,0),c(1,1))
zlm_resid <- do.call(rbind,zlm_cond@hookOut)

# with ngeneson
zlm_cond_sh  <- zlm.SingleCellAssay( ~ condition + cngeneson, sca,  
                                     method='bayesglm', ebayes = TRUE, 
                                     ebayesControl =list(method = "MLE", model = "H1"), 
                                     hook=deviance_residuals_hook )
lrt_cond_sh  <- lrTest(  zlm_cond_sh,CoefficientHypothesis("conditionStim"))
wald_cond_sh <- waldTest(zlm_cond_sh,CoefficientHypothesis("conditionStim"))
logFC_mait   <- getLogFC(zlm_cond_sh,c(1,0,0),c(1,1,0))
zlm_resid_ng <- do.call(rbind,zlm_cond_sh@hookOut)

res_gene          <- data.table(melt(lrt_cond))
res_gene$primerid <- as.character(res_gene$primerid)
res_gene          <- merge(res_gene, fData(sca), by="primerid")
res_gene_hurdle   <- res_gene[metric=="Pr(>Chisq)" & test.type=="hurdle"]
res_gene_hurdle   <- res_gene_hurdle[,adj:=p.adjust(value,"fdr")]
lfc               <- getLogFC(zlm_cond)[contrast=="conditionStim"]
setkey(lfc,primerid)
setkey(res_gene_hurdle,primerid)
res_gene_hurdle   <- merge(lfc,res_gene_hurdle)

res_gene_ng       <- data.table(melt(lrt_cond_sh))
res_gene_ng$primerid <- as.character(res_gene_ng$primerid)
res_gene_ng       <- merge(res_gene_ng, fData(sca), by="primerid")
res_gene_ng_hurdle<- res_gene_ng[metric=="Pr(>Chisq)" & test.type=="hurdle"]
res_gene_ng_hurdle<-res_gene_ng_hurdle[,adj:=p.adjust(value,"fdr")]
lfc_ng               <- getLogFC(zlm_cond_sh)[contrast=="conditionStim"]
setkey(lfc_ng,primerid)
setkey(res_gene_ng_hurdle,primerid)
res_gene_ng_hurdle<-merge(lfc_ng,res_gene_ng_hurdle)

saveRDS( zlm_cond          ,    file = "zlm_cond.rds"   )
saveRDS( zlm_cond_sh       ,    file = "zlm_cond_sh.rds"   )
saveRDS( res_gene_hurdle   ,    file = "res_gene_hurdle.rds"   )
saveRDS( res_gene_ng_hurdle,    file = "res_gene_ng_hurdle.rds"   )

MAST:::layer(sca)  <- 'tpm'
```

# read fit ZLM result
```{r read_saved_data}
 zlm_cond          <-readRDS( file = "../inst/extdata/zlm_cond.rds"   )
 zlm_cond_sh       <-readRDS( file = "../inst/extdata/zlm_cond_sh.rds"   )
 res_gene_hurdle   <-readRDS( file = "../inst/extdata/res_gene_hurdle.rds"   )
 res_gene_ng_hurdle<-readRDS( file = "../inst/extdata/res_gene_ng_hurdle.rds"   )
```

###  The estimated parameters in MAIT

```{r mait_parameters, fig.width=12, fig.height=20  }
genes_signif<- res_gene_ng_hurdle[adj<0.01&abs(logFC)>FCTHRESHOLD][order(logFC)]#[1:100]
genes_notsig<- res_gene_ng_hurdle[adj>0.5 &abs(logFC)<FCTHRESHOLD][order(logFC)]#[1:900]

cc_sig_ng   <- coef(zlm_cond_sh,"C")[c(genes_signif$primerid,genes_notsig$primerid),]
cd_sig_ng   <- coef(zlm_cond_sh,"D")[c(genes_signif$primerid,genes_notsig$primerid),]

par(mfrow=c(6,2))
hist(cc_sig_ng[genes_signif$primerid,   "(Intercept)"]  ,main="signif  cont base")
hist(cc_sig_ng[genes_notsig$primerid,   "(Intercept)"]  ,main="not sig cont base")
hist(cc_sig_ng[genes_signif$primerid,   "conditionStim"],main="signif  cont stim")
hist(cc_sig_ng[genes_notsig$primerid,   "conditionStim"],main="not sig cont stim")
hist(cc_sig_ng[genes_signif$primerid,   "cngeneson"]    ,main="signif  cont cdr ")
hist(cc_sig_ng[genes_notsig$primerid,   "cngeneson"]    ,main="not sig cont cdr ")
hist(cd_sig_ng[genes_signif$primerid,   "(Intercept)"]  ,main="signif  disc base")
hist(cd_sig_ng[genes_notsig$primerid,   "(Intercept)"]  ,main="not sig disc base")
hist(cd_sig_ng[genes_signif$primerid,   "conditionStim"],main="signif  disc stim")
hist(cd_sig_ng[genes_notsig$primerid,   "conditionStim"],main="not sig disc stim")
hist(cd_sig_ng[genes_signif$primerid,   "cngeneson"]    ,main="signif  disc cdr ")
hist(cd_sig_ng[genes_notsig$primerid,   "cngeneson"]    ,main="not sig disc cdr ")
```

###  The estimated stimulation and cdr coefficient in MAIT

```{r mait_cor, fig.width=14, fig.height=7  }
par(mfrow=c(1,2))
plot(zlm_cond_sh@coefD[,"conditionStim"],zlm_cond_sh@coefD[,"cngeneson"], main="D", xlab="stim",ylab="CDR")
plot(zlm_cond_sh@coefC[,"conditionStim"],zlm_cond_sh@coefC[,"cngeneson"], main="C", xlab="stim",ylab="CDR")
```

```{r mait_pair, fig.width=14, fig.height=7 }
par(mfrow=c(1,2))
pairs(data.frame(zlm_cond_sh@coefD[genes_signif$primerid,"conditionStim"],zlm_cond_sh@coefD[genes_signif$primerid,"cngeneson"],zlm_cond_sh@coefC[genes_signif$primerid,"conditionStim"],zlm_cond_sh@coefC[genes_signif$primerid,"cngeneson"]))

par(mfrow=c(1,2))
pairs(data.frame(zlm_cond_sh@coefD[genes_notsig$primerid,"conditionStim"],zlm_cond_sh@coefD[genes_notsig$primerid,"cngeneson"],zlm_cond_sh@coefC[genes_notsig$primerid,"conditionStim"],zlm_cond_sh@coefC[genes_notsig$primerid,"cngeneson"]))
```


```{r set_parameters,eval=FALSE,error=TRUE}
N  <- 200    # number of samples
N0 <- 100    # number of samples in group 0
N1 <- 100    # number of samples in group 1
P  <- 2500   # number of features
P0 <- 100    # number of DE features
P1 <- 2400   # number of not DE features

mait_cdr   <- rowMeans(exprs(sca)[,c(genes_signif$primerid,genes_notsig$primerid)]>0)
mait_cdr_us<- mait_cdr[cData(sca)$condition=="Unstim"]
mait_cdr_st<- mait_cdr[cData(sca)$condition=="Stim"]

ncdr    <- 3
mb_u    <- c(0.2,0.3,0.4)
sb_u    <- 20
mb_s    <- c(0.6,0.5,0.4)
sb_s    <- 20
rho     <- c(0,0.3,0.5,0.8)
nro     <- length(rho)
s       <- c(rep(0,N0),rep(1,N1))

th_c    <- 5
om_c    <- 2

th_d    <- 0
om_d    <- 1

nu_s_c  <- 1
nu_u_c  <- 0
tau_s_c <- 2
tau_u_c <- 1

nu_s_d  <- 3
nu_u_d  <- 0
tau_s_d <- 1 
tau_u_d <- 0.3 

rep_list_no      <-vector("list",nrep)
rep_list_cd      <-vector("list",nrep)
rep_list_limma_no<-vector("list",nrep)
rep_list_limma_cd<-vector("list",nrep)

for(ri in 1:nrep){
  cat("replication: ",ri,"\n")
  cdr_list <-vector("list",ncdr)
  names(cdr_list)<- paste("x",mb_u,mb_s,sep="_")
  for(i in 1:3){
    cdr_list[[i]]<-c(rbeta(N0,mb_u[i]*sb_u,(1-mb_u[i])*sb_u),rbeta(N1,mb_s[i]*sb_s,(1-mb_s[i])*sb_s))
  }

  sigma_d<- runif(P,0,1)
  sdnosh <- log(sqrt(zlm_cond_sh@dispersionNoshrink[c(genes_signif$primerid,genes_notsig$primerid),"C"]))

  varnosh <- zlm_cond_sh@dispersionNoshrink[c(genes_signif$primerid,genes_notsig$primerid),"C"]
  a0=(zlm_cond_sh@priorDOF)/2
  b0=(zlm_cond_sh@priorVar)*a0
  # pdf("dispersion.pdf",width=12,heigh=4)
  #   par(mfrow=c(1,3))
  #   x<-seq(0,20,by=1)
  #   hist(varnosh,probability=T)
  #   curve(dgamma(x,shape=3,rate=0.7),add=TRUE,col="blue")
  #   curve(dinvgamma(x,shape=a0,b0),add=TRUE,col="red")
  #   hist(1/rgamma(100,shape=a0,b0),probability=T,main="zlm estimate")
  #   hist(rgamma(100,shape=3,rate=0.7),probability=T,main="gamma(3,0.7)")
  # dev.off()

  sigma_c <-sqrt( 1/rgamma(P,shape=a0,rate=b0))
  mm_d    <- c( nu_s_d, nu_u_d)[c(rep(1,P0),rep(2,P1))]
  tt_d    <- c(tau_s_d,tau_u_d)[c(rep(1,P0),rep(2,P1))]
  mm_c    <- c( nu_s_c, nu_u_c)[c(rep(1,P0),rep(2,P1))]
  tt_c    <- c(tau_s_c,tau_u_c)[c(rep(1,P0),rep(2,P1))]

  coefs_c<- array(NA,c(P,2,nro))
  coefs_d<- array(NA,c(P,2,nro))
  dimnames(coefs_c)<-list(NULL,c("Stim","CDR"),paste("rho",rho,sep="_"))
  dimnames(coefs_d)<-list(NULL,c("Stim","CDR"),paste("rho",rho,sep="_"))
  for( j in 1:nro){
    Rd <- matrix(c(1,-rho[j],-rho[j],1),2,2)
    Rc <- matrix(c(1,-rho[j],-rho[j],1),2,2)
    for(k in 1:P){
      Dc<-diag(c(tt_c[k],10))
      Dd<-diag(c(tt_d[k],10))
      coefs_d[k,,j]=rmvnorm(1,c(mm_d[k],0),Dd%*%Rd%*%Dd)
      coefs_c[k,,j]=rmvnorm(1,c(mm_c[k],0),Dc%*%Rc%*%Dc)
    }
  }

  stim_d  <- coefs_d[,1,1]
  stim_c  <- coefs_c[,1,1]
  mu_du   <- rnorm( P, th_d, om_d )
  mu_ds   <- (mu_du + stim_d)
  temp    <- mu_du[81:100]
  mu_du[81:100]<- mu_ds[81:100]
  mu_ds[81:100]<- temp
  mu_c    <- rnorm( P, th_c, om_c )
  mu_mat  <- matrix( mu_c, nrow=P, ncol=N )
  stim_mat   <- t(t(matrix(stim_c,nrow=P,ncol=N))*s)
  mustim_mat <- mu_mat+stim_mat
  temp2                     <- mustim_mat[81:100,1:N0 ]
  mustim_mat[81:100,1:N0 ]  <- mustim_mat[81:100,(N0+1):(N0+N1)]
  mustim_mat[81:100,(N0+1):(N0+N1)]<- temp2
  cid    <- matrix(1,P,N)

  disc   <- matrix(NA,P,N)
  dt     <- matrix(NA,P,N)
  colnames(dt)<-paste("cell_",1:N,sep="")

  for(i in 1:P){
    disc[i,]    <- c(rbinom( N0, 1, cid[i,     1:N0     ]*invlogit( mu_du[i] ) ) ,
                     rbinom( N1, 1, cid[i,(N0+1):(N0+N1)]*invlogit( mu_ds[i] ) ) )
  }
  for(i in 1:N){
      dt[,i]    <- disc[,i] * rnorm( P, mustim_mat[,i] , sigma_c ) #sigma_c )
  }

  dt_cd=data.table(mb_u=c(NA,rep(mb_u,nro)),mb_s=c(NA,rep(mb_s,nro)),
            rho=c(NA,rep(rho,each=ncdr)),dt_idx=seq(1,nro*ncdr+1,by=1))
  dt_cd        <- dt_cd[1:4,] # running conditions with no correlation in stimulation and CDR coef
  nsim         <- nrow(dt_cd)
  dt_list      <- vector("list",nrow(dt_cd))
  dt_list[[1]] <- dt
  for(icc in 2:nsim){
    cat("sim: ",icc,"\n")
    te=matrix(NA,P,N)
    colnames(te)<-paste("cell_",1:N,sep="")
    dt_list[[dt_cd[icc,]$dt_idx]]=te
    disc_cd<- matrix(NA,P,N)
    cdr    <- cdr_list[[paste("x",dt_cd$mb_u[icc],dt_cd$mb_s[icc],sep="_")]]
    #for(j in 1:nro){
      rn           <- paste("rho",dt_cd$rho[icc],sep="_")
    stim_d       <- coefs_d[,1,rn]
    stim_c       <- coefs_c[,1,rn]
    beta_cdr_d   <- coefs_d[,2,rn]
    beta_cdr_c   <- coefs_c[,2,rn]
    mu_du        <- rnorm( P, th_d, om_d )
    mu_ds        <- (mu_du + stim_d)
    temp         <- mu_du[81:100]
    mu_du[81:100]<- mu_ds[81:100]
    mu_ds[81:100]<- temp
    mu_c         <- rnorm( P, th_c, om_c )
    mu_mat       <- matrix( mu_c, nrow=P, ncol=N )
    stim_mat     <- t(t(matrix(stim_c,nrow=P,ncol=N))*s)
    mustim_mat   <- mu_mat+stim_mat
    temp2                     <- mustim_mat[81:100,1:N0 ]
    mustim_mat[81:100,1:N0 ]  <- mustim_mat[81:100,(N0+1):(N0+N1)]
    mustim_mat[81:100,(N0+1):(N0+N1)]<- temp2
    for(i in 1:P){
      disc_cd[i,] <- c(rbinom( N0, 1, cid[i,1:N0 ]*invlogit( mu_du[i] + beta_cdr_d[i]*cdr[1:N0] ) ) ,
                       rbinom( N1, 1, cid[i,(N0+1):(N0+N1)]*invlogit( mu_ds[i] + beta_cdr_d[i]*cdr[(N0+1):(N0+N1)]) ) )
    }
    for(i in 1:N){
        dt_list[[dt_cd[icc,]$dt_idx]][,i] <- disc_cd[,i] * rnorm( P, mustim_mat[,i]+beta_cdr_c*cdr[i], sigma_c ) 

    }
    #}
  }

  zlm_sim_no_rho   <- vector( "list", nsim )  
  lrt_sim_no_rho   <- vector( "list", nsim )
  lfc_sim_no_rho   <- vector( "list", nsim )
  zlm_sim_cd_rho   <- vector( "list", nsim )
  lrt_sim_cd_rho   <- vector( "list", nsim )
  lfc_sim_cd_rho   <- vector( "list", nsim )
  fdr_no_rho       <- vector( "list", nsim ) 
  fdr_cd_rho       <- vector( "list", nsim ) 
  fdr_no_fc_rho    <- vector( "list", nsim ) 
  fdr_cd_fc_rho    <- vector( "list", nsim ) 
  limma_no_rho     <- vector( "list", nsim ) 
  limma_cd_rho     <- vector( "list", nsim ) 
  reslimma_no_rho  <- vector( "list", nsim ) 
  reslimma_cd_rho  <- vector( "list", nsim ) 

  neglog10alpha    <- rev(seq(0,500,length.out=1000))
  alpha            <- 10^(-neglog10alpha)
  truth            <- rep(c(0,1),c(P0,P1))
  truth_mat        <- matrix(truth,P,P)
  for(j in 1:nsim){
    cat("creating sca object\n")
    fdat    <- data.frame(primerid=1:P, entrez=1:P, symbolid=1:P, stringsAsFactors=FALSE)
    esetAll <- abind(tpm=t(dt_list[[j]]), rev.along=0)
    cdr_est <- colMeans(dt_list[[j]]>0)
    if( j >= 2 ) cdr_est <- cdr_list[[paste("x",dt_cd$mb_u[j],dt_cd$mb_s[j],sep="_")]]
    cdat    <- data.frame(wellKey=colnames(dt_list[[j]]), condition=factor(s), 
                ncells=1, ngeneson=cdr_est, cngeneson=cdr_est - mean(cdr_est), 
                stringsAsFactors=FALSE)
    rownames(cdat)<-colnames(dt_list[[j]])
    sca_sim     <- FromMatrix('SingleCellAssay', esetAll, cdat, fdat)
    name_vec    <- (1:P)  
    true_count  <- c(P0,P1)
    cat("fitting zlm\n")
    zlm_sim_no_rho[[j]]  <- zlm.SingleCellAssay( ~ condition , sca_sim,  
                                         method='bayesglm', ebayes = TRUE, 
                                         ebayesControl =list(method = "MLE", model = "H1"), 
                                         hook=deviance_residuals_hook )
    lrt_sim_no_rho[[j]]  <- lrTest(  zlm_sim_no_rho[[j]],CoefficientHypothesis("condition1"))
    lfc_sim_no_rho[[j]]  <- getLogFC(zlm_sim_no_rho[[j]],c(1,0),c(1,1))

    zlm_sim_cd_rho[[j]]  <- zlm.SingleCellAssay( ~ condition + cngeneson, sca_sim,  
                                         method='bayesglm', ebayes = TRUE, 
                                         ebayesControl =list(method = "MLE", model = "H1"), 
                                         hook=deviance_residuals_hook )
    lrt_sim_cd_rho[[j]]  <- lrTest(  zlm_sim_cd_rho[[j]],CoefficientHypothesis("condition1"))
    lfc_sim_cd_rho[[j]]  <- getLogFC(zlm_sim_cd_rho[[j]],c(1,0,0),c(1,1,0))
    #saveRDS( zlm_sim_no_rho[[j]]          ,    file = paste("zlm_no",ri,j,".rds",sep="_")   )
    #saveRDS( zlm_sim_cd_rho[[j]]          ,    file = paste("zlm_cd",ri,j,".rds",sep="_")   )

    fdr_no_rho[[j]]    <-sapply(neglog10alpha,function(x)by(-log10(lrt_sim_no_rho[[j]][,"hurdle","Pr(>Chisq)"])>x,truth,sum)/true_count)
    fdr_cd_rho[[j]]    <-sapply(neglog10alpha,function(x)by(-log10(lrt_sim_cd_rho[[j]][,"hurdle","Pr(>Chisq)"])>x,truth,sum)/true_count)
    dt_pos             <- t(exprs(sca_sim))#dt_list[[j]]
    tpm                <- 2^dt_pos
    count_matrix       <- round( t((t(tpm)/colSums(tpm))*1000000))
    rownames(count_matrix)<-name_vec

    cat("fitting limma\n")
    x                <- count_matrix
    new_set          <- ExpressionSet( assayData = as.matrix(x)+1 )

      # without ngeneson
      design           <- model.matrix(~ condition, data=cData(sca_sim))
      colnames(design) <- make.names(colnames(design))
      cont_matrix      <- makeContrasts( "condition1",
                                          levels = design)
      voom.data        <- voom( new_set, design = design )
      voom.data$genes  <- rownames(x)
      voom.fitlimma    <- lmFit( voom.data, design = design )
      voom.fitbayes    <- contrasts.fit(voom.fitlimma,cont_matrix)
      voom.fitbayes    <- eBayes(voom.fitbayes)
      voom.pvalues     <- voom.fitbayes$p.value
      reslimma_no_rho[[j]] <- topTable( voom.fitbayes, adjust.method="fdr", number=Inf, sort.by="none" )

      # with ngeneson
      design_ng           <- model.matrix(~ condition + cngeneson, data=cData(sca_sim))
      colnames(design_ng) <- make.names( colnames(design_ng) )
      cont_matrix_ng      <- makeContrasts( "condition1",
                                             levels = design_ng )
      voom.data_ng        <- voom( new_set, design = design_ng )
      voom.data_ng$genes  <- rownames(x)
      voom.fitlimma_ng    <- lmFit( voom.data_ng, design = design_ng )
      voom.fitbayes_ng    <- contrasts.fit(voom.fitlimma_ng, cont_matrix_ng)
      voom.fitbayes_ng    <- eBayes(voom.fitbayes_ng)
      reslimma_cd_rho[[j]]<- topTable(voom.fitbayes_ng, adjust.method="fdr", number=Inf, sort.by="none" )

    limma_no_rho[[j]]   <- sapply( neglog10alpha, function(x) by(-log10(reslimma_no_rho[[j]]$P.Value) > x, truth, sum )/true_count )
    limma_cd_rho[[j]]   <- sapply( neglog10alpha, function(x) by(-log10(reslimma_cd_rho[[j]]$P.Value) > x, truth, sum )/true_count )
  }
  rep_list_no[[ri]]      <- fdr_no_rho
  rep_list_cd[[ri]]      <- fdr_cd_rho
  rep_list_limma_no[[ri]]<- limma_no_rho
  rep_list_limma_cd[[ri]]<- limma_cd_rho
}

saveRDS( rep_list_no,          file ="rep_list_no.rds"  )
saveRDS( rep_list_cd,          file ="rep_list_cd.rds"  )
saveRDS( rep_list_limma_no,    file ="rep_list_limma_no.rds"  )
saveRDS( rep_list_limma_cd,    file ="rep_list_limma_cd.rds"  )

# dtcdr<-melt(data.frame(abind(cdr_list,along=2),stimulation=rep(c("Unstim","Stim"),each=N0)))
# dtcdr$variable<-factor(dtcdr$variable,labels=c("strong confounding", "moderate confounding", "no confounding"))
# pdf("cdr_confounding.pdf",width=12,heigh=4)
# ggplot(dtcdr)+geom_density()+aes(x=value,color=stimulation)+facet_grid(.~variable)+xlab("CDR") + scale_colour_brewer(palette="Set1")
# dev.off()
```

```{r load_result}

rep_list_no        <- readRDS( file = "../inst/extdata/rep_list_no.rds") 
rep_list_cd        <- readRDS( file = "../inst/extdata/rep_list_cd.rds" ) 
rep_list_limma_no  <- readRDS( file = "../inst/extdata/rep_list_limma_no.rds" ) 
rep_list_limma_cd  <- readRDS( file = "../inst/extdata/rep_list_limma_cd.rds" ) 

# rep_list_no        <- readRDS(file ="rep_list_no.rds"  )
# rep_list_cd        <- readRDS(file ="rep_list_cd.rds"  )
# rep_list_limma_no  <- readRDS(file ="rep_list_limma_no.rds"  )
# rep_list_limma_cd  <- readRDS(file ="rep_list_limma_cd.rds"  )

```
###  ROC 

```{r plot_roc, fig.width=15, fig.height=10}
conditions=c("no CDR in TRUTH", "strong confounding", "moderate confounding", "no confounding")
dttmp=NULL
for(j in 1:4){
dttmp=rbind(dttmp, 
    rbind(data.frame(t(abind(lapply(rep_list_no,function(x)x[[j]]),along=2))[,c(2,1)],
              model="MAST no CDR", condition=conditions[j]),
      data.frame(t(abind(lapply(rep_list_cd,function(x)x[[j]]),along=2))[,c(2,1)],
            model="MAST with CDR", condition=conditions[j]),
      data.frame(t(abind(lapply(rep_list_limma_no,function(x)x[[j]]),along=2))[,c(2,1)],
            model="limma no CDR", condition=conditions[j]),
      data.frame(t(abind(lapply(rep_list_limma_cd,function(x)x[[j]]),along=2))[,c(2,1)],
            model="limma with CDR", condition=conditions[j])
      )
    )
}
names(dttmp)[1:2]<-c("FP","TP")
dttmp$condition = factor(dttmp$condition)
#pdf("simulation_ROC_with_without_CDR.pdf",width=16,height=4)
ggplot(dttmp)+aes(x=FP,y=TP,group=model,color=model)+facet_grid(.~condition)+geom_abline(intercept=0,slope=1,linetype="dotted")+ stat_quantile(method = "rqss",quantiles = c(0.05, 0.95),linetype="dashed", lambda = 1)+ stat_quantile(method = "rqss",quantiles = c(0.5), lambda = 1)+ylim(0,1)+xlim(0,1)
#dev.off()

```